---
title: "Probability distributions and law of large numbers"
subtitle: "EDS 222"
author: "Leonardo Feitosa"
#institute: ""
date: "Fall 2024"
output:
  xaringan::moon_reader:
    self_contained: true
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(leaflet, ggplot2, ggthemes, viridis, dplyr, magrittr, knitr,pagedown,cowplot,latex2exp)
# Define pink color
red_pink <- "#e64173"

# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  # dpi = 300,
  # cache = T,
  warning = F,
  message = F,
  dev = "svg"
)
# A blank theme for ggplot
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_simple <- theme_bw() + theme(
  line = element_blank(),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_blank(),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  # plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
modified_time <- Sys.time() %>% 
  as.numeric() %>% 
  .POSIXct(tz = "UTC") %>% 
  format("%Y%m%d%H%M%OS1")
metathis::meta_tag(metathis::meta(), name = "last_modified", content = modified_time)
```

```{r}
set.seed(123)
```


name: Probability distributions and Law of Large Numbers

#### Moving beyond the normal distribution
- Binomial, Poisson, Log-normal, Uniform...

--

#### Law of large numbers
- Theory behind it with examples


layout: false
class: clear, middle, inverse
# Types of distributions

---

# Types of distributions

## Binomial
> Applies to binary variables
- Stems from repeated Bernoulli trials (count of the number of successful drawls of the expected outcome)
- Outcome corresponds to the 1
- Example: When sampling for the presence of a species in an area, any presence would be computed as a 1 (positive Bernoulli trial).

--

# Types of distributions

## Binomial

```{r, binomial counts, echo = F, out.width = "70%", echo=FALSE, fig.align='center'}
dat <- data.frame(dist = rbinom(n = 1000, size = 20, p = 0.7))

ggplot(data = dat) +
  geom_histogram(aes(x = dist),
               fill = "cyan4") +
  labs(x = "X",
       y = "Count") +
  theme_bw()
```

---

### Probability density function - Binomial distribution

```{r, binomial pdf, echo = F}
dat_pdf <- data.frame(dist = dbinom(0:20, size = 20, p = 0.7))

ggplot(data = dat_pdf) +
  geom_density(aes(x = dist),
               fill = "cyan4") +
  labs(x = "X",
       y = "Probability Density Curve") +
  theme_bw()
```


---

# Types of distributions

## Poisson
> Applies to count data
- Example: How many individuals of a species can we find in an area?
- Differs from the Binomial in the fact that the number of Bernoulli trials is not fixed
- Binomial starts to resemble the Poisson as `n` grows

--

# Types of distributions

## Poisson

```{r, exponential counts, echo = F}
dat_poi <- data.frame(dist = rpois(n = 1000, lambda = 14))

ggplot(data = dat_poi) +
  geom_histogram(aes(x = dist),
                 color = "black",
                 fill = "cyan4") +
  labs(x = "X",
       y = "Count") +
  theme_bw()
```

---

### Probability density function - Poisson distribution

```{r, poisson pdf, echo = F}
poi_pdf <- data.frame(dist = dpois(0:30, lambda = 14))

ggplot(data = poi_pdf) +
  geom_density(aes(x = dist),
               fill = "cyan4") +
  labs(x = "X",
       y = "Probability Density Curve") +
  theme_bw()
```


---

# Types of distributions

## Are the Poisson and Binomial distributions discrete or continuous?

--

# Types of distributions

## Exponential
> Applies to continuous data

--

# Types of distributions

## Exponential

```{r, exponential dist, echo = F}
exp_data <- data.frame(dist = rexp(n = 10000, rate = 0.2))

ggplot(data = exp_data) +
  geom_histogram(aes(x = dist),
                 fill = "cyan4",
                 color = "black") +
  labs(x = "",
       y = "Frequency") +
  theme_bw()
```

---

### Probability density function - Exponential distribution

```{r, exponential pdf, echo = F}
exp_pdf <- data.frame(dist = dexp(0:50, rate = 0.2))

ggplot(data = exp_pdf) +
  geom_density(aes(x = dist),
                 fill = "cyan4",
                 color = "black") +
  labs(x = "X",
       y = "Frequency") +
  theme_bw()
```

--

# Types of distributions

## Lognormal
> Applies to continuous data

--

## Probability density function - Lognormal distribution
```{r, lognormal, echo = F}
set.seed(1298)
samp <- as.data.frame(rlnorm(1000))
  colnames(samp) = "x"
samp <- samp %>% mutate(x = round(x,digits = 2))

# Plot it
ggplot(data = samp, aes(x=x)) +
  geom_density(fill="cyan4") +
  theme_bw() 
  
```

--

# Law of large numbers

> The mean obtained from a large number of independent random samples will converge to the true value, assuming it exists in reality. 

```{r, lln, echo = F}
set.seed(1234)
flips <- data.frame(flips = sample(c("heads", "tails"), size = 1000, replace = TRUE),
                    n_flips = c(1:1000))

ggplot(data = flips) +
  geom_line(aes(x = n_flips, y = cumsum(flips == "heads")/(1:length(flips)))) +
  geom_hline(yintercept = 0.5,
             color = "cyan4") +
  labs(x = "Number of coin flips",
       y = "Proportion of heads") +
  theme_bw()
```

---

class: center, middle


Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

Some slide components were borrowed from [The Stat 311 Development Team](https://uw-statistics.github.io/Stat311Tutorial/limit-theorems.html) awesome teaching materials.

---

exclude: true

```{R, printpdfs, echo = F, eval = F}
pagedown::chrome_print(
  input = "02-summstats.html",
  output = "02-summstats.pdf",
  timeout = 90
)
```
















